---
title: "Chapter 5"
author: "Amanda P"
date: "1/17/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 5.3.1 Exercises

1. The first two arguments to `ggplot` are `data` and `mapping`. The first two arguments to all layer functions are `mapping` and `data`. Why does the order of teh arguments differ? (Hint: think about what you set most commonly.)

2. The following code uses dplyr to generate some summary statistics about each class of car (you'll learn how it works in Chapter 10). 

```{r}
library(dplyr)
class <- mpg %>% group_by(class) %>% summarise(n=n(), hwy=mean(hwy))
class
```

Use the data to recreate the plot (shown in book).

```{r}
ggplot(mpg,aes(class,hwy)) + geom_point()
#TODO: ADD MORE TO THIS PLOT
```

# 5.4.3 Exercises

1. Simplify the following plot spcifications:


```{r}

```

```{r}

```

```{r}

```


2. What does the following code do? Does it work? Does it make sense? Why/why not?

```{r}
ggplot(mpg) + geom_point(aes(class,cty)) + geom_boxplot(aes(trans,hwy))
```

3. What happens if you try to use a continuous variable on the x axis in one layer, and a categorical variable in another layer? What happens if you do it in the opposite order?


# 5.5.1 Exercises

1. Download and print out the ggplot2 cheatsheet from http://www.rstudio.com/resources/cheatsheets/ so you have a handy visual reference for all the geoms.

Done check.

2. Look at the documentation for the graphical primitive geoms. Which aesthetics do they use? How can you summarize them in a compact form?

3. What's the best way to master an unfamiliar geom? List three resources to help you get started.

4. For each of the plots below, identify the geom used to draw it.

5. For each of the following problems, suggest a useful geom:

* Display how a variable has changed over time
* Show the detailed distribution of a single variable
* Focus attention on the overall trend in a large dataset
* Draw a map
* Label outlying points

# 5.6.2 Exercises

1. The code below creates a similar dataset to `stat_smooth()`. Use the appropriate geoms to mimic the default `geom_smooth()` display.

```{r}
mod <- loess(hwy ~ displ, data=mpg)
smoothed <- data.frame(displ= seq(1.6,7,length=50))
pred <- predict(mod, newdata=smoothed, se=TRUE)
smoothed$hwy <- pred$fit
smoothed$hwy_lwr <- pred$fit - 1.96 * pred$se.fit
smoothed$hwy_upr <- pred$fit + 1.96 * pred$se.fit
```

2. What stats were used to create the following plots?

* First plot:
* Second plot:
* Third plot:

3. Read the help for `stat_sum()` then `geom_count()` to create a plot tha tshows the proportion of cars that have each combination of `drv` and `trans`.

#5.7.1 Exercises

1. When might you use `position_nudge()`? Read the documentation.

2. Many position adjustments can only be used with a few geoms. For example, you can't stack boxplots or eror bars. Why not? What properties must a geom possess in order to be stackable? What properties must it possess to be dodgeable?

3. Why might you use `geom_jitter()` instead of `geom_count()`? What are teh advantages and disadvantages of each technique?

4. When might you use a stacked area plot? What are the advantages and disadvantages compared to a line plot?

