---
title: "Chapter 5"
author: "Amanda P"
date: "1/17/2018"
output:   
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

# 5.3.1 Exercises

1. The first two arguments to `ggplot` are `data` and `mapping`. The first two arguments to all layer functions are `mapping` and `data`. Why does the order of the arguments differ? (Hint: think about what you set most commonly.)

Most often we define the dataset in the `ggplot` function, which means that we most often will not need to define the data in the layer functions. The mapping is sometimes determined in `ggplot` and sometimes in the layer function.

2. The following code uses dplyr to generate some summary statistics about each class of car (you'll learn how it works in Chapter 10). 

```{r}
library(dplyr)
class <- mpg %>% group_by(class) %>% summarise(n=n(), hwy=mean(hwy))
class
```

Use the data to recreate the plot (shown in book).

```{r}
ggplot(mpg,aes(class,hwy)) + geom_jitter(height=.25,width=.075) + geom_point(aes(class,hwy), data=class, color="red", size=4) + geom_text(aes(x=class,y=10,label=sprintf("n=%d",n)),data=class,)
```

# 5.4.3 Exercises

1. Simplify the following plot spcifications:

```{r}
ggplot(mpg) + geom_point(aes(displ,hwy))
```

```{r}
ggplot(mpg,aes(hwy,cty)) + geom_point() + geom_smooth()
```

The following code was given in the book:

```{r}
ggplot(diamonds,aes(carat,price)) + geom_point(aes(log(brainwt),log(bodywt)),data=msleep)
```

Take out the diamonds data since we don't use it:

```{r}
ggplot(msleep, aes(log(brainwt),log(bodywt))) + geom_point()
```

2. What does the following code do? Does it work? Does it make sense? Why/why not?

```{r}
ggplot(mpg) + geom_point(aes(class,cty)) + geom_boxplot(aes(trans,hwy))
```

The code above works. It plots `class` vs `cty` using points and on the same plot `trans` vs `hwy`. It doesn't make sense. Issues: 1) the y label only shows `cty` so it's not clear that `hwy` is also plotted, and 2) it doesn't make alot of sense to use both points and boxplots on the same plot.

3. What happens if you try to use a continuous variable on the x axis in one layer, and a categorical variable in another layer? What happens if you do it in the opposite order?

First, we'll try a continuous variable on the x-axis (`hwy`) in one layer, and a categorical variable (`class`) in another layer.

The code below returns the following error:
`Error: Discrete value supplied to continuous scale`

`#ggplot(mpg) + geom_point(aes(hwy,cty)) + geom_point(aes(class,cty))`

Next, we'll reverse the order of the two layers such that the categorical variable is plotted on the x-axis first and then the continuous variable. This order works.

```{r}
ggplot(mpg) + geom_point(aes(class,cty)) + geom_point(aes(hwy,cty))
```


# 5.5.1 Exercises

1. Download and print out the ggplot2 cheatsheet from http://www.rstudio.com/resources/cheatsheets/ so you have a handy visual reference for all the geoms.

Done.

2. Look at the documentation for the graphical primitive geoms. Which aesthetics do they use? How can you summarize them in a compact form?

The graphical primitive geoms use the following aesthetics.

* `geom_point()`: x, y, alpha, color, fill, group, shape, size, stroke
* `geom_path()`: x, y, alpha, color, group, linetype, size
* `geom_ribbon()`: x, ymin, xmax, alpha, color, fill, group, linetype, size 
* `geom_segment()`: x, y, xend, yend, alpha, color, group, linetype, size
* `geom_rect()`: xmin, xmax, ymin, ymax, alpha, color, fill, group, linetype, size
* `geom_polyon()`: x, y, alpha, color, fill, group, linetype, size
* `geom_text()`: x, y, label, alpha, angle, color, family, fontface, group, hjust, lineheight, size, vjust

Summarize them in a compact form:
+ position: x,y, ymin, ymax, xmin, xmax, angle, hjust, vjust
+ style: alpha, color, fill, shape, size, stroke, linetype, fontface, lineheight


3. What's the best way to master an unfamiliar geom? List three resources to help you get started.

* Best way: Read the help documentation using `?geom_()`
* Google examples.
* Try it!

4. For each of the plots below, identify the geom used to draw it.

Reading left to right:

* `geom_violin`
* `geom_point`
* `geom_hex` (Note that package `hexbin` is required.)
* `geom_jitter`
* `geom_line`
* `geom_path`

5. For each of the following problems, suggest a useful geom:

* Display how a variable has changed over time: `geom_line`
* Show the detailed distribution of a single variable: `geom_bar` (discrete), `geom_density` (continuous)
* Focus attention on the overall trend in a large dataset: `geom_smooth`
* Draw a map: `geom_map`
* Label outlying points: `geom_text`

# 5.6.2 Exercises

1. The code below creates a similar dataset to `stat_smooth()`. Use the appropriate geoms to mimic the default `geom_smooth()` display.

```{r}
mod <- loess(hwy ~ displ, data=mpg)
smoothed <- data.frame(displ= seq(1.6,7,length=50))
pred <- predict(mod, newdata=smoothed, se=TRUE)
smoothed$hwy <- pred$fit
smoothed$hwy_lwr <- pred$fit - 1.96 * pred$se.fit
smoothed$hwy_upr <- pred$fit + 1.96 * pred$se.fit

ggplot(mpg,aes(displ,hwy)) + geom_point()  + geom_ribbon(aes(x=displ,ymin=hwy_lwr,ymax=hwy_upr),data=smoothed,fill="grey70") + geom_line(aes(displ,hwy),data=smoothed,alpha=.7)
```

2. What stats were used to create the following plots?

* First plot: `stat_ecdf`
* Second plot: `stat_qq` 
* Third plot: `stat_bin`

3. Read the help for `stat_sum()` then `geom_count()` to create a plot that shows the proportion of cars that have each combination of `drv` and `trans`.

```{r}
ggplot(mpg,aes(drv,trans)) + geom_count(aes(size = ..prop.., group=1))
```



#5.7.1 Exercises

1. When might you use `position_nudge()`? Read the documentation.

`position_nudge()` is especially useful with text to place it in a position where it is readable and not hiding data points.

2. Many position adjustments can only be used with a few geoms. For example, you can't stack boxplots or error bars. Why not? What properties must a geom possess in order to be stackable? What properties must it possess to be dodgeable?

* You can't stack boxplots or error bars. Why not?

Boxplots and error bars describe the variability of a single variable. Only one such description per variable makes sense.

* What properties must a geom possess in order to be stackable? 
 
 The geom can be used to describe different segments of the variable of interest. For example, a bar chart (bars can be stacked) can be used to describe a variable such as frequency of an income category using stacked bars to differentiate between male and female. 

* What properties must it possess to be dodgeable? Ans: TODO

Similar to stackable geoms, dodgeable geoms must have multiple values per variable. We could use dodging in the example above to compare salaries of males to salaries of females to place the bars side by side. This would be more useful to compare the heights of the bars.  

3. Why might you use `geom_jitter()` instead of `geom_count()`? What are the advantages and disadvantages of each technique?

* `geom_jitter()`: Might use this for smaller datasets. Con: For large datasets, jitter may not be enough to see all points.
* `geom_count()`: Might use this if the data size is larger. Size of point indicates count or proportion of datapoints at that location. Con?

4. When might you use a stacked area plot? What are the advantages and disadvantages compared to a line plot?

* stacked area plot: Use when you are interested in trends of combinations of variables. The $n^{th}$ area shows the trend of the combination of the first $n$ variables.
* line plot: Use when you are interested in comparing the trend of each variable separately (and in contrast to one another). 

