---
title: "Chapter 10"
author: "Amanda P"
date: "2/11/2018"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
```

# 10.2.3 Exercises

1. Practice your filtering skills by:

Finding all the diamonds with equal x and y dimensions

```{r}
filter(diamonds,x==y)
```

A depth between 55 and 70:

```{r}
filter(diamonds,depth >= 55, depth <= 70)
```

A carat smaller than the median carat:

```{r}
filter(diamonds,carat < median(carat))
```

Cost more than $10,000 per carat:

```{r}
filter(diamonds, round(price/carat, digits=2) > 10000)
```

Are of good or better quality:

```{r}
filter(diamonds, cut >= "Good")
```

2. Fill in the question marks in the table:

Given table:

------------------------------
Expression | TRUE | FALSE | NA
-----------|------|-------|---
x          | X    |       |
?          |      | X     |    
is.na(x)   |      |       | X
!is.na(x)  | ?    | ?     | ?
?          | X    |       | X
?          |      | X     | X

Solution:

------------------------------
Expression | TRUE | FALSE | NA
-----------|------|-------|---
x          | X    |       |
!x         |      | X     |    
is.na(x)   |      |       | X
!is.na(x)  | ?    | ?     | ?
?          | X    |       | X
?          |      | X     | X

TODO: Finish above table. I'm not exactly sure what it's asking. Is x the same expression from row to row?

3. Repeat the analysis of outlying values with $z$. Compared to $x$ and $y$, how would you characterize the relationship of $x$ and $z$, or $y$ and $z$?

Analysis of $x$ and $z$:

```{r}
ggplot(diamonds, aes(x,z)) + geom_bin2d()
```

Next filter out points where $x$ or $z$ are 0. Also, filter out large $z$ values (greater than 30). 

```{r}
diamondsFiltered <- filter(diamonds, x>0, z>0, z<30)
ggplot(diamondsFiltered, aes(x,z)) + geom_bin2d()
```

We can see in the above plot that $x$ and $z$ have a linear relationship.

Next consider $y$ and $z$:

```{r}
ggplot(diamonds, aes(y,z)) + geom_bin2d()
```

Let's filter out outliers (y>30 or z>30) and remove zeros:

```{r}
diamondsFiltered <- filter(diamonds, z<30, y<30, z>0, y>0)
ggplot(diamondsFiltered, aes(y,z)) + geom_bin2d()
```

We see in the plot above that $y$ and $z$ also have a linear relationship.

4. Install the `ggplot2movies` package and look at the movies that have a missing budget. How are they different from the movies with a budget? (Hint: try a frequency polygon plus coor = is.na(budget))

```{r}
#install.packages("ggplot2movies")
library("ggplot2movies")
filter(movies,is.na(budget))
```

Compare ratings:

```{r}
ggplot(movies, aes(rating, ..density.., color=is.na(budget))) + geom_freqpoly()
```

5. What is `NA & FALSE` and `NA | TRUE`? Why? Why doesn't `NA * 0` equal zero? What number time zero does not equal 0? What do you expect NA ^ 0 to equal? Why?

```{r}
NA & FALSE

NA | TRUE
```

FALSE "and-ed" with anything is always FALSE. So, `NA & FALSE` is FALSE. TRUE "or-ed" with anything is always TRUE. So, `NA | TRUE` is TRUE.

`NA * 0` does not equal zero, because `Inf * 0` does not equal 0. Since we don't know what the missing value (NA) is, we can't say that its product with 0 is 0.

`NA ^ 0` is 1. 


#10.3.2 Exercises 

1. Practice your variable creation skills by creating the following new variables:

* The approximate volume of the diamond (using x, y, and z).

```{r}
diamonds2 <- mutate(diamonds, volume = x*y*z)
select(diamonds2,x,y,z,volume) %>% head()
```

* The approximate density of the diamond.

```{r}
#Density is mass/volume.
diamonds2 <- mutate(diamonds2, density=carat/volume)
select(diamonds2,carat,volume,density) %>% head()
```

* The price per carat.

```{r}
diamonds2 <- mutate(diamonds2,ppc=round(price/carat, digits=2))
select(diamonds2,price,carat,ppc) %>% head()
```

* Log transformation of carat and price.

```{r}
diamonds2 <- mutate(diamonds2,caratL=log(carat),priceL=log(price))
select(diamonds2,carat,caratL,price,priceL) %>% head()
```


2. How can you improve the data density of `ggplot(diamonds, aes(x,z)) + stat_bin2d()`? What transformation makes it easier to extract outliers?

```{r}
ggplot(diamonds, aes(log(x),log(z))) + stat_bin2d()
```

3. The depth variable is just the width of the diamond (average of x and y) divided by its height (z) multiplied by 100 and rounded to the nearest integer. Compute the depth yourself and compare it to the existing depth variable. Summarize your findings with a plot.

NOTE: Diamonds documention says that "depth percentage = z / mean(x, y)", which is the reciprocal of the description in this problem.  

```{r}
diamondsNew <- mutate(diamonds2,depth2=round(100*z*2/(x+y),1))
select(diamondsNew,x,y,z,depth2,depth) %>% head()
ggplot(diamondsNew, aes(depth,depth2)) + geom_point()
```

4. Compare the distribution of symmetry for diamonds with $x > y$ vs. $x < y$.

```{r}
diamonds2 <- mutate(diamonds2,sym=x-y)
select(diamonds2,x,y,sym) %>% head()
#Filter outliers using filter, then plot:
ggplot(filter(diamonds2,sym>-1,sym<1), aes(sym,color=x<y)) + geom_freqpoly()
```

# 10.4.3 Exercises

1. For each year in the `ggplot2movies::movies` data determine the percent of movies with missing budgets. Visualize the result.

```{r}
missingBudget <- movies %>% group_by(year) %>% summarize(pct=mean(is.na(budget)))
ggplot(missingBudget, aes(year,pct)) + geom_line()
```

2. How does the average length of a movie change over time? Display your answer with a plot, including some display of uncertainty.


```{r}
lenByDecade <- movies %>% mutate(decade=floor(movies$year/10)*10) %>% group_by(decade) %>% summarize(avgLen=mean(length),se=sd(length))
head(lenByDecade)

ggplot(lenByDecade,aes(decade,avgLen,ymin=avgLen-se,ymax=avgLen+se)) + geom_point() + geom_line() + geom_errorbar()
```


3. For each combination of diamond quality (e.g. cut, color, and clarity), count the number of diamonds, the average price and the average size. Visualize the results.

```{r}
diamonds2 <- diamonds %>% mutate(size=sqrt(x^2 + y^2)) %>% group_by(cut,color,clarity) %>% summarize(cnt=n(),avgPrice=mean(price),avgSize=mean(size)) 

ggplot(diamonds2, aes(avgSize,avgPrice, color=color)) + geom_point(aes(size=cnt)) + facet_grid(cut ~ clarity) + xlab("Average Size") + ylab("Average Price")
```


4. Compute a histogram of carat by "hand" using a binwidth of 0.1. Display the results with `geom_bar(stat = "identity")`. (Hint: you might need to create a new variable first.)

```{r}
diamonds2 <- diamonds %>% mutate(bin=round(carat,digits=1))
histData <- diamonds2 %>% group_by(bin) %>% summarize(n=n())

ggplot(histData, aes(bin,n)) + geom_bar(stat="identity") + xlab("carat") + ylab("count")
```


5. In the baseball example, the batting average seems to increase as the number of at bats increases. Why?

Recall ba = H/AB.

It's likely that as batters are at bat more (and bat against more pitchers) their performance improves.

# 10.5.1 Exercises

1. Translate each of the examples in this chapter to use the pipe. 

Page 212:

```{r}
diamonds %>% group_by(clarity) %>% summarize(price=mean(price))
```

```{r}
diamonds %>% group_by(cut,depth) %>% summarize(n=n()) %>% filter(depth>55,depth<70)
```

Page 216:

```{r}
#install.packages("Lahman")
library("Lahman")
Batting %>% filter(AB>0) %>% group_by(playerID) %>% summarize(ba=sum(H, na.rm=TRUE) / sum(AB, ra.rm=TRUE))
```


2. What does the following pipe do?

```{r}
library(magrittr)
x <- runif(10)
x %>%
  subtract(mean(.)) %>%
  raise_to_power(2) %>%
  mean() %>%
  sqrt()
```

Subtract the mean(x) from each x value, then square each value, then find the mean of all values, then take the square root. 


3. Which player in the `Batting` dataset has had the most consistently good performance over the course of their career?

Of all batters who have atleast 80 at bats, cobbty01 has the highest batting average (0.36636) with 4189 hits out of 11,434 total at bats (over the course of 24 years).

```{r}
baPerBatter <- Batting %>% 
    group_by(playerID) %>% summarize(totalHits=sum(H, na.rm=TRUE),totalAB=sum(AB, na.rm=TRUE),ba=sum(H, na.rm=TRUE)/sum(AB, na.rm=TRUE),n=n()) %>% 
  filter(totalHits>0,totalAB>50) %>% 
  arrange(desc(ba))

head(baPerBatter)

```

Additionally, we can see that his batting average did not vary too greatly over the years (especially after the improvement during the first 5 years).

```{r}
cobbty01 <- Batting %>% filter(playerID=="cobbty01") %>% 
  mutate(ba=H/AB)

ggplot(cobbty01, aes(yearID,ba,ymin=0,ymax=0.5)) + geom_line()

```


