---
title: "Chapter 9"
author: "Amanda P"
date: "2/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(devtools)
library(tidyr)
library(dplyr)
```

# 9.3.3 Exercises

1. How can you translate each of the initial example datasets into the other form?

TODO: WHich datasets? How to get ec2 dataset?


2. How can you convert back and forth between the `economics` and `econometrics_long` datasets built into ggplot2?

Economics to economics_long:

```{r}
gather(economics,key="variable",value="value","pce","pop","psavert","uempmed","unemploy")
```

TODO: What is value01 in economics_long?

Economics_long to economics:

```{r}
spread(select(economics_long,1:3),key=variable,value=value)
```


3. Install the EDAWR package from https://github.com/rstudio/EDAWR. Tidy the `storms`, `population` and `tb` datasets.

```{r}
#install_github("rstudio/EDAWR")
library(EDAWR)
```

```{r}
storms
gather(storms,key="variable",value="value","wind","pressure")
```

TODO: The following spread command is throwing errors when knitting document. However, it runs OK on the command line.

```{r}
head(population)
#spread(population,key=country,value=population)
```

```{r}
tb
gather(tb,key="ageGroup",value="count","child","adult","elderly")
```

# 9.4.1

1. Install the EDAWR package from https://githug.com/rstudio/EDAWR. Tidy the `who` dataset.

```{r}
head(who)
gather(who,key="new",value="value",5:60)
```

2. Work through the demos included in the `tidyr` package (`demo(package = "tidyr")`).

```{r}
demo('so-15668870')
```

```{r}
demo('so-16032858')
```

```{r}
demo('so-17481212')
```

```{r}
demo('so-9684671')
```



