---
title: "Chapter 11"
author: "Amanda P"
date: "2/24/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(magrittr)
library(dplyr)
```

# 11.2.1 Exercises

1. What happens if you repeat the above analysis with all diamonds? (Not just all diamonds with two or fewer carats). What does the strange geometry of `log(carat)` vs relative price represent? What does the diagonal line without any points represent?

Consider the linear relationship between `lcarat` and `lprice`:

```{r}
diamonds2 <- diamonds %>% 
  mutate(
    lcarat = log2(carat),
    lprice = log2(price)
  )
diamonds2

ggplot(diamonds2, aes(lcarat,lprice)) +
  geom_bin2d() +
  geom_smooth(method="lm",se=FALSE,size=2,color="yellow")
```
We can see above that the x-axis has been extended to the right compared to the original plot (since the original plot was limited to 2 or fewer carats). The range of `lprice` did not increase. 


```{r}
mod <- lm(lprice ~ lcarat, data=diamonds2)
diamonds2 <- diamonds2 %>% mutate(rel_price = resid(mod))
diamonds2
ggplot(diamonds2,aes(carat,rel_price)) + geom_bin2d()
```

The above plot is not good as it is showing a distinct pattern in the residuals. The linear model is overpredicting the log price for large carats. The variance of the residuals is significantly higher for smaller carats. Looking back at the picture of the linear model, the model extends beyond the range of `lprice`, thus there are only points below the line at this point (which is why we see negative residual.) 

```{r}
color_cut <- diamonds2 %>%
  group_by(color, cut) %>%
  summarize(
    price = mean(price),
    rel_price = mean(rel_price)
  )
color_cut
```


```{r}
ggplot(color_cut, aes(color, price)) +
  geom_line(aes(group=cut), color="grey80") + 
  geom_point(aes(color = cut))
```

Above we see the same trend as in the book, where the lowest quality diamonds (color=J) hae the highest price, especially for Premium cut diamonds. Plotting relative price against color solves the issue.

```{r}
ggplot(color_cut, aes(color, rel_price)) +
  geom_line(aes(group = cut), color = "grey80") + 
  geom_point(aes(color=cut))
```

2. I made an unsupported assertion that lower-quality diamonds tend to be larger. Support my claim with a plot.

```{r}
ggplot(diamonds, aes(color, carat)) + geom_boxplot()

```

3. Can you create a plot that simultaneously shows the effect of color, cut, and clarity on relative price? If there's too much information to show on one plot, think about how you might create a sequence of plots to convey the same message.

```{r}
color_cut_clarity <- diamonds2 %>%
  group_by(color, cut, clarity) %>%
  summarize(
    price = mean(price),
    rel_price = mean(rel_price)
  )
ggplot(color_cut_clarity, aes(color,rel_price)) + 
  geom_line(aes(group=cut), color="grey80") + 
  geom_point(aes(color=cut)) + 
  facet_wrap(~clarity)
```

4. How do depth and table related to the relative price?

Table and depth do not have much relatiohship with relative price. This makes sense as table and depth are highly correlated to carat (weight) and we've already taken out the effect of carat.

```{r}
#diamonds2
ggplot(diamonds2,aes(depth,rel_price)) +
  geom_bin2d()
```

```{r}
#diamonds2
ggplot(diamonds2,aes(table,rel_price)) +
  geom_bin2d()
```

# 11.3.1 Exercises

1. The final plot shows a lot of short-term noise in the overall trend. How could you smooth this further to focus on long-term changes?

Instead of using `geom_line`, we can draw a loess curve using `geom_smooth`:

```{r}
txhousing

#de-season function:
deseas <- function(x,month){
  resid(lm(x ~ factor(month), na.action = na.exclude))
}

txhousing <- txhousing %>%
  group_by(city) %>%
  mutate(rel_sales = deseas(log(sales), month))

ggplot(txhousing, aes(date,rel_sales)) + 
  geom_line(aes(group=city), alpha = 1/5) +
  #geom_line(stat = "summary", fun.y="mean", color="red")
  geom_smooth(method="loess",se=FALSE)
```

2. If you look closely (e.g. `+ xlim(2008,2012)`) at the long-term trend you'll notice a weird pattern in 2009-2011. It looks like there was a big dip in 2010. Is this dip "real"? (i.e. can you spot it in the original data)

Looking at the relative sales (after taking out the month effect), a number of cities do indeed tend to have a dip in 2010:

```{r}
ggplot(txhousing, aes(date,rel_sales)) + 
  geom_line(aes(group=city), alpha = 1/5) +
  geom_line(stat = "summary", fun.y="mean", color="red") +
  #geom_smooth(method="loess",se=FALSE) +
  xlim(2008,2012)
```

TODO: Maybe there are more missing data points in 2010? Check.

```{r}
sales2010 <- txhousing %>% filter(year==2010)
sales2010
```


3. What other variables in the TX housing data show strong seasonal effects? Does this technique help to remove them?

```{r}
summary(txhousing)

#"Months inventory": amount of time it would take to sell all current listings at current pace of sales. 
ggplot(txhousing, aes(date,log(inventory))) + 
  geom_line(aes(group=city), alpha=1/2)

#Total active listings
ggplot(txhousing, aes(date,log(listings))) +
  geom_line(aes(group=city), alpha=1/2)

#Total value of sales
ggplot(txhousing, aes(date,log(volume))) +
  geom_line(aes(group=city), alpha=1/2)

#Median sale price:
ggplot(txhousing, aes(date,log(median))) +
  geom_line(aes(group=city), alpha=1/2)
```


4. Not all the cities in this data set have complete time series. Use your `dplyr` skills to figure out how much data each city is missing. Display the results with a visualization.

```{r}
numDates <- txhousing %>% 
  summarize(nDates=n_distinct(date)) %>% 
  select(nDates) %>% 
  max()
numDates

cityCnts <- txhousing %>%
  na.omit() %>%
  group_by(city) %>%
  summarize(nComplete=n()) %>%
  mutate(pctComplete=nComplete/numDates)
cityCnts

ggplot(cityCnts, aes(city,pctComplete)) + 
  geom_col() +
  labs(x="City", y="% Complete") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

  
```


5. Replicate the computation that `stat_summary()` did with dplyr so you can plot the data "by hand".

Comment: I think this question is refering to the `geom_line` shown on page 229 using `stat="summary"` and `fun.y="mean"`.

```{r}
meanRelSalesData <- txhousing %>%
  group_by(date) %>%
  summarize(meanRelSales = mean(rel_sales, na.rm=TRUE))
meanRelSalesData

ggplot(txhousing, aes(date,rel_sales)) + 
  geom_line(aes(group=city), alpha = 1/5) +
  geom_line(aes(date,meanRelSales),data=meanRelSalesData,color="red")

```

# 11.5.1 Exercises

1. Do your conclusions change if you use a different measurement of model fit like AIC or deviance? Why/ why not?

```{r}
#install.packages("broom")
library(broom)
```

```{r}
models <- txhousing %>%
  group_by(city) %>%
  do(mod = lm(log2(sales) ~ factor(month), data = ., na.action = na.exclude))
models

model_sum <- models %>% glance(mod)
model_sum

ggplot(model_sum, aes(AIC, reorder(city,AIC))) +
  geom_point()
```

Looking at AIC: Harlingen, San Marcos, and Brownsville have the highest AIC (2 of the three are the same cities that had the lowest $R^2$). Dallas, El Paso, and Midland have the lowest AIC, which are different than those having the highest $R^2$. (Recall low AIC is better.)

```{r}
worstAIC <- c("Harlingen", "San Marcos", "Brownsville")
bestAIC <- c("Dallas", "El Paso", "Midland")
extreme <- txhousing %>% ungroup() %>%
  filter(city %in% c(worstAIC,bestAIC), !is.na(sales)) %>%
  mutate(city = factor(city, c(worstAIC, bestAIC)))

ggplot(extreme, aes(month, log(sales))) +
  geom_line(aes(group=year)) +
  facet_wrap(~city)
```

TODO: Why the change?


2. One possible hypothesis that explains why McAllen, Harlingen and Brownsville have lower $R^2$ is that they're smaller towns so there are fewer sales and more noise. Confirm or refute this hypothesis.

Looking at the plot on page 234, McAllen (low $R^2$) and Bryan-College Station (high $R^2$) have similar order of magnitude of sales. This is also apparent looking at a boxplot of their sales. 

```{r}
txhousing %>% ungroup() %>%
  group_by(city) %>%
  filter(city %in% c("McAllen","Bryan-College Station"), !is.na(sales)) %>%
  summarize(meanSales=mean(sales))

selectCities <- txhousing %>%
  filter(city %in% c("McAllen","Bryan-College Station"), !is.na(sales)) 

ggplot(selectCities, aes(city,sales)) + geom_boxplot()

```

If we look at all of the extreme cities, however, we do see that NE Tarrant County (high $R^2$) does have significantly more sales. Also, Brownsville and Harlengen (both having low $R^2$) do have fewer sales than the others.

```{r}
selectCities <- txhousing %>%
  filter(city %in% c("McAllen","Bryan-College Station", "Lubbock", "NE Tarrant County", "Brownsville", "Harlingen"), !is.na(sales)) 

ggplot(selectCities, aes(city,sales)) + geom_boxplot()
```


3. McAllen, Harlingen and Brownsville seem to have much more year-to-year variation that Bryan-College Station, Lubbock, and NE Tarrant County. How does the model change if you also include a linear trend for year? (i.e. `log(sales) ~ factor(month) + year`).

```{r}
models <- txhousing %>%
  group_by(city) %>%
  do(mod = lm(log2(sales) ~ factor(month) + year, data = ., na.action = na.exclude))
models

model_sum <- models %>% glance(mod)
model_sum

ggplot(model_sum, aes(r.squared, reorder(city,r.squared))) +
  geom_point()

```

Harlingen moved up 10 places (comparing the $R^2$ values) by adding year to the model. 

4. Create a faceted plot that shows the seasonal patterns for all cities. Order the facets by the $R^2$ for the city.

```{r}
models <- txhousing %>%
  group_by(city) %>%
  do(mod = lm(log2(sales) ~ factor(month), data = ., na.action = na.exclude))

model_sum <- models %>% glance(mod)
model_sum

ggplot(txhousing, aes(month, log(sales))) +
  geom_line(aes(group=year)) +
  facet_wrap(~city)
```

TODO: How to order facets by $R^2$? Maybe join datasets and then us reorder statment.

#11.6.1 Exercises

1. Pull out the three cities with the highest and lowest seasonal effect. Plot their coefficients. 

```{r}
coefs <- models %>% tidy(mod)
coefs

#Pull out months:
months <- coefs %>%
  filter(grepl("factor",term)) %>%
  tidyr::extract(term,"month","(\\d+)", convert=TRUE)
months

#Strength of seasonal effect:
coefSummary <- months %>%
  group_by(city) %>%
  summarize(max=max(estimate))
#ggplot(coefSummary, aes(2^max, reorder(city,max))) +
#  geom_point() + 
#  ggtitle("Maximum coefficient per City")

#Extract highest and lowest seasonal effect:
top3 <- coefSummary %>% top_n(3,max)
bottom3 <- coefSummary %>% top_n(-3,max)
selectCitySummary <- rbind(top3,bottom3)

#Plot their coeficients:
ggplot(selectCitySummary, aes(2^max, reorder(city,max))) +
  geom_point() + 
  ggtitle("Maximum 2^coefficients for Cities \nwith highest and lowest seasonal effect")
```


2. How does strength of seasonal effect relate to the $R^2$ for the model? Answer with a plot.

There seems to be a positive relationship between $R^2$ and seasonal effect:

```{r}
#model_sum contains r.squared for each city 
#coefSummary contains max coef (seasonal effect)

combined <- inner_join(model_sum,coefSummary,by="city")
combined

ggplot(combined, aes(r.squared,2^max)) + 
  geom_point() +
  labs(y="Max 2^Coef (Seasonal effect)")
```


3. You should be extra cautious when your results agree with your prior beliefs. How can you confirm or refute my hypothesis about the causes of strong seasonal patterns?

Hypothesis: Cities with strongest seasonal effect are college and beach towns.
TODO: Finish

4. Group the diamonds data by cut, clarity, and color. Fit a linear model `log(price) ~ log(carat)`. What does the intercept tell you? What does the slope tell you? How do the slope and intercept vary across the groups? Answer with a plot.

The slope tells us how log(carat) changes in relationship with changes in log(price) (for each cut, color, and quality). The intercept tells us the starting log(price) reguarless of log(carat).

```{r}
models <- diamonds %>%
  group_by(cut,clarity,color) %>%
  do(mod = lm(
    log(price) ~ log(carat),
    data = .,
    na.action = na.exclude
  ))
models

models %>% glance(mod)

models %>% tidy(mod)

```

TODO: How to plot each model?

# 11.7.1 Exercises

1. A common diagnostic plot is fitted values (`.fitted`) vs. residuals (`.resid`). Do you see any patterns? What if you include the city or month on the same plot?

```{r}

```


2. Create a time series of `log(sales)` for each city. Highlight points tha thave a standardized residual of greater than 2.

