---
title: "Chapter 9"
author: "Amanda P"
date: "2/7/2018"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(devtools)
library(tidyr)
library(dplyr)
library(EDAWR) #needed for 9.3.3
```

# 9.3.3 Exercises

1. How can you translate each of the initial example datasets into the other form?

TODO: WHich datasets? How to get ec2 dataset?


2. How can you convert back and forth between the `economics` and `econometrics_long` datasets built into ggplot2?

Economics to economics_long:

```{r}
gather(economics,key="variable",value="value","pce","pop","psavert","uempmed","unemploy")
```


Economics_long to economics:

```{r}
spread(select(economics_long,1:3),key=variable,value=value)
```


3. Install the EDAWR package from https://github.com/rstudio/EDAWR. Tidy the `storms`, `population` and `tb` datasets.

```{r}
#install_github("rstudio/EDAWR")
#library(EDAWR)
```

The storms data is already tidy. We'll gather the population and tb datasets.

```{r}
head(EDAWR::population)
gather(EDAWR::population,key="year",value="population",`1995`:`2013`)
```


```{r}
tb
gather(tb,key="ageGroup",value="count","child":"elderly")

#Alternative method (from @joonwoob). This might be easier to read:
#tb %>% gather(child:elderly, key = "ageGroup", value = "count")
```

# 9.4.1

1. Install the EDAWR package from https://githug.com/rstudio/EDAWR. Tidy the `who` dataset.

```{r}
head(who)
gather(who,key="new",value="value",5:60)
```

2. Work through the demos included in the `tidyr` package (`demo(package = "tidyr")`).

```{r}
demo('so-15668870')
```

```{r}
demo('so-16032858')
```

```{r}
demo('so-17481212')
```

```{r}
demo('so-9684671')
```



