---
title: "Chapter8"
author: "Amanda P"
date: "2/3/2018"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

# 8.2.1 Exercises

1. Try out all the themes in ggthemes. Which do you like the best?

```{r}
install.packages("ggthemes")
library(ggthemes)
```



2. What aspects of the default theme do you like? What don't you like? What would you change?

First define base graphic: 

```{r}
df <- data.frame(x=1:3, y=1:3)
base <- ggplot(df, aes(x,y)) + geom_point()
```

The following themes are from the `ggtheme` package and not already shown in the book.

* theme_base: a theme resembling the default base graphics in R. See also theme_par.

```{r}
base + theme_base() + ggtitle("theme_base()")
```


* theme_calc: a theme based on LibreOffice Calc.

```{r}
base + theme_calc() + ggtitle("theme_calc()")
```

* theme_economist: a theme based on the plots in the The Economist magazine.

```{r}
base + theme_economist() + ggtitle("theme_economist()")
```


* theme_few: theme from Stephen Few's "Practical Rules for Using Color in Charts".

```{r}
base + theme_few() + ggtitle("theme_few()")
```

* theme_fivethirtyeight: a theme based on the plots at fivethirtyeight.com.

```{r}
base + theme_fivethirtyeight() + ggtitle("theme_fivethirtyeight()")
```

* theme_gdocs: a theme based on Google Docs.

```{r}
base + theme_gdocs() + ggtitle("theme_gdocs()")
```

theme_hc: a theme based on Highcharts JS.

```{r}
base + theme_hc() + ggtitle("theme_hc()")
```

* theme_par: a theme that uses the current values of the base graphics parameters in par.

```{r}
base + theme_par() + ggtitle("theme_par()")
```

* theme_pander: a theme to use with the pander package.

```{r}
base + theme_pander() + ggtitle("theme_pander()")
```


* theme_stata: themes based on Stata graph schemes.

```{r}
base + theme_stata() + ggtitle("theme_stata()")
```



* theme_wsj: a theme based on the plots in the The Wall Street Journal.

```{r}
base + theme_wsj() + ggtitle("theme_wsj()")
```


3. Look at the plots in your favorite scientific journal. What themes do they most resemble? What are the main differences?

TODO: Finish


# 8.4.6 Exercises

1. Create the ugliest plot possible! (Contributed by Andrew D. Steen, University of Tenessee - Knoxville)

```{r}
base + ggtitle("Ugly Plot") + theme(plot.background = element_rect(fill="red"),panel.background = element_rect(fill="lightgreen"),panel.grid.major.x = element_line(color="gray60"),plot.title= element_text(face="italic",color="yellow",angle = 10),axis.title = element_text(color="purple", angle = -30 )) 
```


2. `theme_dark()` makes the inside of the plot dark, but not the outside. Change the plot background to black, and then update the text settings so you can still read the labels.

```{r}
base + theme_dark() + theme(plot.background = element_rect(fill="black"),axis.text = element_text(color="white"),axis.title = element_text(color="white"))
```

3. Make an elegant theme that uses "linen" as the background color and a serif font for the text.

```{r}
base + ggtitle("This is a plot") + theme(panel.background = element_rect(fill="linen"),plot.title= element_text(family="serif"),axis.title = element_text(family="serif"))
```


4. Systematically explore the effects of `hjust` when you have a multiline title. Why doesn't `vjust` do anything?

Left justified:

```{r}
base + ggtitle("This is a very very very very very\nlooooooooooooong plot title") + theme(plot.title = element_text(hjust=0))
```

Centered:

```{r}
base + ggtitle("This is a very very very very very\nlooooooooooooong plot title") + theme(plot.title = element_text(hjust=.5))
```

Right justified:

```{r}
base + ggtitle("This is a very very very very very\nlooooooooooooong plot title") + theme(plot.title = element_text(hjust=1))
```

Printed over the edge of the plot: 

```{r}
base + ggtitle("This is a very very very very very\nlooooooooooooong plot title") + theme(plot.title = element_text(hjust=1.5))
```

TODO: The question is "Why doesn't vjust do anything?" However, `vjust` seems to move the text vertically:

```{r}
base + ggtitle("This is a very very very very very\nlooooooooooooong plot title") + theme(plot.title = element_text(vjust=4))

```


