---
title: "Chapter 10"
author: "Amanda P"
date: "2/11/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
```

# 10.2.3 Exercises

1. Practice your filtering skills by:

Finding all the diamonds with equal x and y dimensions

```{r}
filter(diamonds,x==y)
```

A depth between 55 and 70:

```{r}
filter(diamonds,depth >= 55, depth <= 70)
```

A carat smaller than the median carat:

```{r}
filter(diamonds,carat < median(carat))
```

Cost more than $10,000 per carat:

```{r}
filter(diamonds, price > 10000)
```

Are of good or better quality:

```{r}
filter(diamonds, cut >= "Good")
```

2. Fill in the question marks in the table:

------------------------------
Expression | TRUE | FALSE | NA
-----------|------|-------|---
x          | X    |       |
!x         |      | X     |    
is.na(x)   |      |       | X
!is.na(x)  | ?    | ?     | ?
?          | X    |       | X
?          |      | X     | X

TODO: Finish above table. I'm not exactly sure what it's asking. Is x the same expression from row to row?

3. Repeat the analysis of outlying values with $z$. Compared to $x$ and $y$, how would you characterize the relationship of $x$ and $z$, or $y$ and $z$?

Analysis of $x$ and $z$:

```{r}
ggplot(diamonds, aes(x,z)) + geom_bin2d()
```

Next filter out points where $x$ or $z$ are 0. Also, filter out large $z$ values (greater than 30). 

```{r}
diamondsFiltered <- filter(diamonds, x>0, z>0, z<30)
ggplot(diamondsFiltered, aes(x,z)) + geom_bin2d()
```

We can see in the above plot that $x$ and $z$ have a linear relationship.

Next consider $y$ and $z$:

```{r}
ggplot(diamonds, aes(y,z)) + geom_bin2d()
```

Let's filter out outliers (y>30 or z>30) and remove zeros:

```{r}
diamondsFiltered <- filter(diamonds, z<30, y<30, z>0, y>0)
ggplot(diamondsFiltered, aes(y,z)) + geom_bin2d()
```

We see in the plot above that $y$ and $z$ also have a linear relationship.

4. Install the `ggplot2movies` package and look at the movies tha thave a missing budget. How are they different from the movies with a budget? (Hint: try a frequency polygon plus coor = is.na(budget))

```{r}
#install.packages("ggplot2movies")
library("ggplot2movies")
filter(movies,is.na(budget))
```

Compare ratings:

```{r}
ggplot(movies, aes(rating, ..density.., color=is.na(budget))) + geom_freqpoly()
```

5. What is `NA & FALSE` and `NA | TRUE`? Why? Why doesn't `NA * 0` equal zero? What number time zero does not equal 0? What do you expect NA ^ 0 to equal? Why?

```{r}
NA & FALSE

NA | TRUE
```

FALSE "and-ed" with anything is always FALSE. So, `NA & FALSE` is FALSE. TRUE "or-ed" with anything is always TRUE. So, `NA | TRUE` is TRUE.

`NA * 0` does not equal zero, because `Inf * 0` does not equal 0. Since we don't know what the missing value (NA) is, we can't say that its product with 0 is 0.

`NA ^ 0` is 1. 

